---
title: 'Selected Topics of Psychometrics (NMST570)'
subtitle: 'Homework Assignment 6: DIF'
author: 'Kateřina Hudáčová (Benjamín Kunc)'
date: \today
output:
  html_document:
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float: yes
    fig_caption: yes
    df_print: paged
    code_folding: show
---

Note: The assignment is due on December 20, 2021. Send your code and html/pdf report to `dlouha@cs.cas.cz` (and to `martinkova@cs.cas.cz` in copy).

</br>

#### 1. Read and comment on Chapter 8 on https://perusall.com. [2 pts]

#### 2. Run the code for the Chapter 8, available on course webpage http://www.cs.cas.cz/martinkova/NMST570.html.

#### 3. Training section in ShinyItemAnalysis [3 pts]
##### Use ShinyItemAnalysis interactive application, and finish Exercises in DIF/Fairness/Exercises.

#### 4. Project preparation.
##### Explore the Czech matura data OR DATA OF YOUR CHOICE as described below. You can work with 1-2 teammates on this task. All teammate names have to be included on your answer sheet. [5 pts]

I am working with the `BFI2` dataset from the `ShinyItemAnalysis` package. The `BFI2` dataset contains responses of $1733$ Czech respondents to the Big Five Inventory 2, as well as information about their age, education, and gender. The BFI2 items are scored using a 5-point Likert scale.

```{r load_data}
library(deltaPlotR)
library(difNLR)
library(difR)
library(ggplot2)
library(ltm)
library(mirt)
library(ShinyItemAnalysis)
library(cowplot)
library(Cairo)
library(dplyr)

data(BFI2, package='ShinyItemAnalysis')

# Data frame for agreeableness domain
BFI2_a <- BFI2[, seq(2, 57, 5)]
BFI2_a <- cbind(BFI2_a, BFI2[61:63])
BFI2_a <- BFI2_a %>% 
  mutate(total_score = rowSums(BFI2_a[1:12]))
```

I created a data frame for the *agreeability* domain, which I'm going to be working with in this assignment.

</br>

#### 4a. Introduction to DIF
##### Explain why DIF analysis is important in test validation. Introduce your data. Discuss which variables could you use for testing for DIF, provide their descriptive statistics.

DIF analysis is important in test validation, because it takes into account that groups may differ in overall ability and allows us to examine potential differences on the level of individual items. It can uncover unfair items or in case of tests that are meant to function differently for different groups, confirm their instructional sensitivity.
</br>

There are three variables that could technically be used for DIF testing -- gender, age (converted to age groups) and education.

```{r DIFintro}
# Gender
BFI2_a$Gender <- factor(BFI2_a$Gender, levels = c(0,1), labels = c("Female", "Male"))
table(BFI2_a$Gender)
tapply(BFI2_a$total_score, BFI2_a$Gender, summary)

# Age
summary(BFI2_a$Age)
BFI2_a$Age <- factor(dplyr::case_when(BFI2_a$Age <= 20 ~ '15-20',
                 TRUE ~ '21-26'))
table(BFI2_a$Age)
tapply(BFI2_a$total_score, BFI2_a$Age, summary)

# Education
BFI2_a$Educ <- factor(dplyr::case_when(BFI2_a$Educ <= 4 ~ "Secondary",
                                      BFI2_a$Educ > 4 ~ "Tertiary"))
                               
summary(BFI2_a$Educ)
tapply(BFI2_a$total_score, BFI2_a$Educ, summary)
```
There are $1003$ women and $730$ men in the `BFI2´ dataset. As can be seen in the summary statistic, the total agreeability scores seem slightly different for male and female participants with female participants scoring higher in agreeability (\mu = $44.3$) than their male counterparts (\mu = $41.95$).
</br>

When it comes to age, the ´BFI2´ sample is rather homogeneous with all participants being 15 to 26 years old. I divided participants into two groups -- 15-20 (N = $1055$) and 21-26 (N = $678$) years old. The summary statistic shows, that the observed agreeability is a bit different for age groups with younger participants scoring (on average) lower (\mu = $42.82$) than older participants (\mu = $44.07$).
</br>

The `Educ` variable has quite a lot of missing values (NA = $1010$). Nonetheless, I divided participants into two groups based on education level (secondary or tertiary). The observed ability in both groups is, again, slightly different with participants with tertiary education scoring higher (\mu = $44.64$) than participants with secondary education (\mu = $41.85$).

#### 4b. DIF analysis with Delta Plot
##### Binarize your data. Using the Delta plot method, which items are detected as functioning differently when using a fixed threshold? Which items are detected as functioning differently when using a threshold based on normal approximation? What is the value of the threshold now?

I binarized my data and tried computing Delta plots for all the above mentioned grouping variables.

```{r DeltaPlot}
# Binarize data
bin <- function(x) {
  ifelse(x %in% 4:5, 1, 0)
}

BFI2_a_bin <- data.frame(lapply(BFI2_a[1:12], bin))

# Gender
BFI2_a_binG <- BFI2_a_bin %>% 
  mutate(Gender = BFI2_a$Gender)


# Delta plot with fixed threshold
(DP_fixed <- deltaPlotR::deltaPlot(data = BFI2_a_binG, group = "Gender",
                                   focal.name = "Female", thr = 1.5))


# Delta plot with normal approximation threshold
(DP_norm <- deltaPlotR::deltaPlot(data = BFI2_a_binG, group = "Gender",
                                  focal.name = "Female", thr = "norm"))

# Age
BFI2_a_binA <- BFI2_a_bin %>% 
  mutate(Age = BFI2_a$Age)

# Delta plot with fixed threshold
(DP_fixed <- deltaPlotR::deltaPlot(data = BFI2_a_binA, group = "Age",
                                   focal.name = "15-20", thr = 1.5))
# Delta plot with normal approximation threshold
(DP_norm <- deltaPlotR::deltaPlot(data = BFI2_a_binA, group = "Age",
                                  focal.name = "15-20", thr = "norm"))
deltaPlotR::diagPlot(DP_norm, thr.draw = TRUE, main ="DP_norm Age")

# Education
BFI2_a_binE <- BFI2_a_bin %>% 
  mutate(Educ = BFI2_a$Educ)

# Delta plot with fixed threshold
(DP_fixed <- deltaPlotR::deltaPlot(data = BFI2_a_binE, group = "Educ",
                                   focal.name = "Secondary", thr = 1.5))
deltaPlotR::diagPlot(DP_fixed, thr.draw = TRUE, main = "DP_fixed Education")

# Delta plot with normal approximation threshold
(DP_norm <- deltaPlotR::deltaPlot(data = BFI2_a_binE, group = "Educ",
                                  focal.name = "Secondary", thr = "norm"))
deltaPlotR::diagPlot(DP_norm, thr.draw = TRUE, main = "DP_norm Education")

```

While using a fixed threshold and education as a grouping variable, item 6 was detected as functioning differently. When using a threshold based on normal approximation, item 6 was detected as functioning differently with a threshold of $1.2715$. Item 6 refers to item 27 in the complete inventory, and is phrased as: *"Považuji se za někoho, kdo má v povaze odpouštět."*. This item is also detected when using a normal approximation and age as a grouping variable with a threshold of $0.6756$. 

</br>

#### 4c. DIF analysis with Maentel Haenszel test
##### Use the binarized data. Using the MH test, which items are detected as functioning differently? What are the values of DIF effect sizes for DIF items? What is the odds ratio for item 1, interpret the result. Discuss any items detected as DIF, provide wording or explanation for DIF.

I used the `difMH` function from the `difR` package and gender as a grouping variable.

```{r MH}
difR::difMH(BFI2_a_binG, group = "Gender", focal.name = "Female")
```

Using the Mantel-Haenszel test, the following items were detected as functioning differently when grouping by gender: i2, i7, i17, i27, i32. The effect size for item i2 is large ($2.0368$), the effect for item i7 is moderate ($-1.0765$) and the rest is negligible. On average, the odds of agreeing with item i2 (*"Považuji se za někoho, kdo je soucitný, má dobré srdce."*) are about two times higher for females. On average, the odds of with item i7 (*"Považuji se za někoho, kdo je uctivý, s ostatními zachází s úctou."*) are higher for males. Both effects could be presumably explained by gender roles and corresponding societal expectations.

</br>

#### 4d DIF analysis with regression models
##### Select a regression model for DIF detection based on your data. Using the regression method, which items are detected as functioning differently? Provide interpretations: Which items favor one and which the other group? Is there any item performing non-uniform DIF? How do the results change when using Benjamini-Hochberg correction for multiple comparison? Provide table of item parameters and interpret for at least one DIF item. Choose some DIF items and provide item information curves. Provide predictionf for correct answer from members of one and of the other group.

Since I'm working with ordinal data, I selected a cumulative logit model.

```{r DIFregr}
(fit <- difNLR::difORD(Data = BFI2_a[1:12], group = BFI2_a$Gender, focal.name = "Female", model = "cumulative"))

coef(fit, SE = TRUE)$i2
coef(fit, SE = TRUE)$i17
coef(fit, SE = TRUE)$i42

plot(fit, item = "i2", plot.type = "cumulative", group.names = c("Females", "Males"))
plot(fit, item = "i2", plot.type = "category", group.names = c("Females", "Males"))

predict(fit, item = "i2", match = 0, group = c(0, 1))
predict(fit, item = "i2", match = 0, group = c(0, 1), type = "cumulative")

(fit_correc <- difNLR::difORD(Data = BFI2_a[1:12], group = BFI2_a$Gender, focal.name = "Female", model = "cumulative", p.adjust.method = "BH"))
```

Items 2, 7, 17, 42 a 47 were detected as functioning differently based on gender. In order to at least agree a little with item 2 (*"Považuji se za někoho, kdo je soucitný, má dobré srdce."*) with a 50% probability, male respondents were estimated to have a standardized total score of $-1.16$, while a somewhat lower standardized total score ($-1.86$) was sufficient for female respondents. A similar trend can be observed with item 17 (reversed - *"Považuji se za někoho, kdo s ostatními příliš nesoucítí.*), which makes sense given the wording of the two items. Item 42 on the other hand slightly favors male respondents, who, in order to at least agree a little bit with the statement *"Považuji se za někoho, kdo je vůči záměrům ostatních nedůvěřivý."* (reversed) with a 50% probability, were estimated to have a standardized total score of $1.08$. Female respondents were estimated to have a slightly higher total score of $1.15$. A male respondent with an average agreeability score has a 60% probability of seeing himself at least somewhat as someone who is sympathetic, while a female respondent has a probability of 63%. With the Bejamini-Hochberg correction, the detected items remain the same, but the p-value of item 42 is now only below the threshold $0.001$.

</br>


#### 4e. DIF analysis with IRT models
##### Fit a group specific model with mirt() function. Which items are detected as DIF? Provide interpretations.

I fit a 2PL model for both genders.

```{r DIFirt}
(fit_irt0 <- difR::itemParEst(BFI2_a_binG[BFI2_a_binG$Gender == "Female", 1:12], model = "2PL"))
(fit_irt1 <- difR::itemParEst(BFI2_a_binG[BFI2_a_binG$Gender == "Male", 1:12], model = "2PL"))
```

As we can see, discrimination and difficulty of item 2 are very different for males and females. Most items were easier for females, item 42 and 47 are an exception.

</br>

#### 4f. (extra) Interactive ShinyItemAnalysis app.
##### Use the interactive ShinyItemAnalysis application to create some of the plots or to create a report with selected DIF methods. If your data is too large, consider using sample data (e.g. from a random sample of 2000 respondents and selection of items - single domain, etc.).

Your text (description of methods).

</br>


#### 5. Use/discuss your own data and/or provide constructive feedback to the lecture, exercises, or any of the material. [BONUS 1 pt]

Your text.

</br>
